Surrogate model for an aligned-spin effective one body waveform model of binary neutron star inspirals using Gaussian process regression

### 2018-12-02 paper/paper_no_event.tex
Paper: changes in the abstract

Fast and accurate waveform models are necessary for measuring the properties of inspiraling binary neutron star systems such as GW170817. We present a completed version of the aligned-spin binary neutron star waveform model using the effective one body formalism known as \texttt{SEOBNRv4T}. This model includes the quadrupolar and octopolar adiabatic and dynamical tides, and we improve it by including the spin-induced quadrupole moment effect and a prescription for tapering the end of the waveform to agree with numerical simulations. The resulting model has 14 intrinsic parameters, so we use universal relations that approximate all matter effects in terms of the leading quadrupolar tidal parameters. Unfortunately, the current implementation can take up to an hour to evaluate, and this is too slow for many parameter estimation codes that require $\mathcal{O}(10^7)$ sequential waveform evaluations. We therefore construct a fast, frequency-domain surrogate of this model using Gaussian process regression. The resulting surrogate has a maximum mismatch of $4.5\times 10^{-4}$ for the Advanced LIGO detector, and requires $\sim 0.13$~s to evaluate for a waveform with a starting frequency of 20~Hz.

>>>>

Fast and accurate waveform models are necessary for measuring the properties of inspiraling binary neutron star systems such as GW170817. We present a frequency-domain surrogate version of the aligned-spin binary neutron star waveform model using the effective one body formalism known as \texttt{SEOBNRv4T}. This model includes the quadrupolar and octopolar adiabatic and dynamical tides. The version presented here is improved by the inclusion of the spin-induced quadrupole moment effect, and completed by a prescription for tapering the end of the waveform to reproduce qualitatively numerical simulations. The resulting model has 14 intrinsic parameters, so we reduce its dimensionality by using universal relations that approximate all matter effects in terms of the leading quadrupolar tidal parameters. The previous implementation of this model could take up to an hour to evaluate, and this is too slow for many parameter estimation codes that require $\mathcal{O}(10^7)$ sequential waveform evaluations. We therefore construct a fast and faithful frequency-domain surrogate of this model using Gaussian process regression. The resulting surrogate has a maximum mismatch of $4.5\times 10^{-4}$ for the Advanced LIGO detector, and requires $\sim 0.13$~s to evaluate for a waveform with a starting frequency of 20~Hz.
